---
title: "Bachelor Thesis - Ania Safronova, 2023 *** *****"
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
library(forcats)
library(stringi)
library(readxl)
library(plm)
library(sandwich)
library(stargazer)
library(ExPanDaR)
```

# Data preprocessing

```{r}
data = read_xlsx('thesis_data.xlsx')

data$Right_to_Vote = as.numeric(data$Right_to_Vote)
data$Female_Years_School = as.numeric(data$Female_Years_School)
data$Educ_Gender_Equality = as.numeric(data$Educ_Gender_Equality)
data$Investments = as.numeric(data$Investments)
data$n = rowSums(!is.na(data)) - 2

#соединим датасет категорий стран (по income и region) с рабочим датасетом
class = read_xlsx('data.xlsx', sheet = 'class')
class = class[3:220,]
colnames(class) = c('Country', 'Region', 'Income_Group')
class$Income_Class = case_when(class$Income_Group == "Low income" ~ 1,
                               class$Income_Group == "Lower middle income" ~ 2,
                               class$Income_Group == "Upper middle income" ~ 3,
                               class$Income_Group == "High income" ~ 4)
 
data = data %>% full_join(class, by = 'Country')
data = data %>% filter(Year <= 2000)
```

# Descriptive statistics

```{r}
data2 = data %>% select(-Year, -n, -Income_Class)
t = prepare_descriptive_table(data2)
t$kable_ret
```

# Outliers

```{r}
Q1 = quantile(data$Educ_Gender_Equality, .25, na.rm = TRUE)
Q3 = quantile(data$Educ_Gender_Equality, .75, na.rm = TRUE)
IQR = IQR(data$Educ_Gender_Equality, na.rm = TRUE)
subset(data, data$Educ_Gender_Equality < (Q1 - 1.5*IQR) | data$Educ_Gender_Equality > (Q3 + 1.5*IQR)) %>% select(Country, Year, Educ_Gender_Equality)

#Gabon is an obvious outlier (Figure 1)

list = c()
for (i in 1:length(data$Country)){
  list = rbind(list, ifelse(data$Country[i] == 'Gabon', 'Gabon', ''))
}
gabons = data.frame(gabons = c(list))
ggplot(data, aes(Country, Educ_Gender_Equality, color = Region)) + geom_point() + theme_minimal() + ylab('educational gender equality') + scale_color_manual(values =  c("#f7db86", "#ad2431", "#f4eed3", "#9ab19a", '#533549', '#7a6c5d', '#202647')) + xlab('') + theme(panel.background = element_rect(fill = 'white', color = 'white'), panel.grid.major = element_line(color = 'grey', size = 0.02)) + geom_text(label = gabons$gabons, nudge_x = 14, check_overlap = T, size = 3) + theme(axis.text.x=element_blank())



Q1 = quantile(data$Investments, .25, na.rm = TRUE)
Q3 = quantile(data$Investments, .75, na.rm = TRUE)
IQR = IQR(data$Investments, na.rm = TRUE)
subset(data, data$Investments < (Q1 - 1.5*IQR) | data$Investments > (Q3 + 1.5*IQR)) %>% select(Country, Year, Investments)

#Zimbabwe is an obvious outlier (Figure 2)

list = c()
for (i in 1:length(data$Country)){
  list = rbind(list, ifelse(data$Country[i] == 'Zimbabwe', 'Zimbabwe', ''))
}
zimbabves = data.frame(zimbabves = c(list))
ggplot(data, aes(Country, Investments, color = Region)) + geom_point() + theme_minimal() + ylab('investments in education') + scale_color_manual(values =  c("#f7db86", "#ad2431", "#f4eed3", "#9ab19a", '#533549', '#7a6c5d', '#202647')) + xlab('') + xlim(data$Country) + theme(panel.background = element_rect(fill = 'white', color = 'white'), panel.grid.major = element_line(color = 'grey', size = 0.02)) + geom_text(label = zimbabves$zimbabves, nudge_x = -700, check_overlap = T, size = 3) + theme(axis.text.x = element_blank()) 
```

# Show that the dataset is unbalanced

Calculate the frequency of observations of each variable
And then the average frequency for all variables for each country (row_mean)

```{r}
n0 = data %>% dplyr::select(Year, Country, Right_to_Vote) %>% drop_na(Year)
n0$n = rowSums(!is.na(n0)) - 2
n0 = n0 %>% group_by(Country) %>% summarise(Right_to_Vote_frequency = sum(n)/length(n))

n1 = data %>% dplyr::select(Year, Country, Share_in_Parliament) %>% drop_na(Year)
n1$n = rowSums(!is.na(n1)) - 2
n1 = n1 %>% group_by(Country) %>% summarise(Share_in_Parliament_frequency = sum(n)/length(n))

n2 = data %>% dplyr::select(Year, Country, GDP) %>% drop_na(Year)
n2$n = rowSums(!is.na(n2)) - 2
n2 = n2 %>% group_by(Country) %>% summarise(GDP_frequency = sum(n)/length(n))

n3 = data %>% dplyr::select(Year, Country, Total_Fertility_Rate) %>% drop_na(Year)
n3$n = rowSums(!is.na(n3)) - 2
n3 = n3 %>% group_by(Country) %>% summarise(Total_Fertility_Rate_frequency = sum(n)/length(n))

n4 = data %>% dplyr::select(Year, Country, Democracy_Estimate) %>% drop_na(Year)
n4$n = rowSums(!is.na(n4)) - 2
n4 = n4 %>% group_by(Country) %>% summarise(Democracy_Estimate_frequency = sum(n)/length(n))

n5 = data %>% dplyr::select(Year, Country, Female_Years_School) %>% drop_na(Year)
n5$n = rowSums(!is.na(n5)) - 2
n5 = n5 %>% group_by(Country) %>% summarise(Female_Years_School_frequency = sum(n)/length(n))

n6 = data %>% dplyr::select(Year, Country, Educ_Gender_Equality) %>% drop_na(Year)
n6$n = rowSums(!is.na(n6)) - 2
n6 = n6 %>% group_by(Country) %>% summarise(Educ_Gender_Equality_frequency = sum(n)/length(n))

n7 = data %>% dplyr::select(Year, Country, Investments) %>% drop_na(Year)
n7$n = rowSums(!is.na(n7)) - 2
n7 = n7 %>% group_by(Country) %>% summarise(Investments_frequency = sum(n)/length(n))

n8 = data %>% dplyr::select(Year, Country, Anti_Patriarchy_Score) %>% drop_na(Year)
n8$n = rowSums(!is.na(n8)) - 2
n8 = n8 %>% group_by(Country) %>% summarise(Anti_Patriarchy_Score_frequency = sum(n)/length(n))

#frequency of observations for each variable
observations = Reduce(function (...) { merge(..., all = TRUE) }, list(n0, n1, n2, n3, n4, n5, n6, n7, n8))
                      
#how many observations on average for all variables
observations$row_mean = rowMeans(observations[ , 2:10])

observations = observations %>% full_join(class, by = 'Country')
```

Dependence of the number of observations on the region of the country

```{r}
observations2 = observations %>% drop_na(Region)

## Figure 3
ggplot(observations2, aes(x = Region, y = row_mean, color = Region)) + geom_boxplot() + ylab('mean frequency of observations among variables') + xlab('') + geom_jitter(position = position_jitter(0.2)) + scale_color_manual(values =  c("#f7db86", "#ad2431", "#f4eed3", "#9ab19a", '#533549', '#7a6c5d', '#202647')) + theme_minimal() + theme(axis.text.x = element_text(angle = 15, vjust = 0.8, hjust = 0.75, size = 8, color = 'black'), legend.position = "none", plot.title = element_text(hjust = 0.5)) + ggtitle('Countries by Region')
```

Dependence of the number of observations on the welfare of the country

```{r}
observations1 = observations %>% drop_na(Income_Group)

## Figure 4
ggplot(observations1, aes(x = Income_Group, y = row_mean, color = Income_Group)) + geom_boxplot() + ylab('mean frequency of observations among variables') + geom_jitter(position = position_jitter(0.2)) + scale_color_manual(values =  c("#f7db86", '#202647', "#ad2431", "#9ab19a")) + theme_minimal() + theme(axis.text.x = element_text(color = 'black', size = 9.5), legend.position = "none", plot.title = element_text(hjust = 0.5)) + aes(x = fct_inorder(Income_Group)) + xlab('') + ggtitle('Countries by Income Level')
```

# Connection between the regressors and the dependent variable

Dependence of Share_in_Parliament from GDP for different groups of countries

```{r}
# Scatter plot (Figure 5)

exper4 = data %>% drop_na(GDP, Share_in_Parliament) %>% group_by(Region, Year) %>% summarise(gdp = median(GDP), share = median(Share_in_Parliament)) %>% drop_na(Region)
ggplot() + geom_point(aes(exper4$gdp, exper4$share, col = as.factor(exper4$Region))) + theme_minimal() + xlab('median GDP') + ylab('median share of women in parliament') + scale_colour_manual(name = "Region") + scale_colour_manual(name = "Region", values = c("#f7db86", "#ad2431", "#f4eed3", "#9ab19a", '#533549', '#7a6c5d', '#202647'))


# Detalization

## Figure 6
europe = data %>% filter(Region == 'Europe & Central Asia')
model = lm(GDP ~ Share_in_Parliament, europe)
ggplot(europe) + geom_point(aes(Share_in_Parliament, GDP), col = '#ad2431') + theme_minimal() + xlab('share in parliament') + ylab("GDP") + ggtitle('Europe & Central Asia') + theme(plot.title = element_text(hjust = 0.5)) + geom_abline(slope = coef(model)[["Share_in_Parliament"]], intercept = coef(model)[["(Intercept)"]], size = 0.8)

## Figure 7
america = data %>% filter(Region == 'North America')
model = lm(GDP ~ Share_in_Parliament, america)
ggplot(america) + geom_point(aes(Share_in_Parliament, GDP), col = '#533549') + theme_minimal() + xlab('share in parliament') + ylab("GDP") + ggtitle('North America') + theme(plot.title = element_text(hjust = 0.5)) + geom_abline(slope = coef(model)[["Share_in_Parliament"]], intercept = coef(model)[["(Intercept)"]], size = 0.6)
```

Dependence of Right_to_Vote from GDP for different groups of countries (by region)

```{r}
# Europe & Central Asia, right to vote (Figure 8)

europe = data %>% filter(Region == 'Europe & Central Asia')
europe$Right_to_Vote = as.factor(europe$Right_to_Vote)
europe = europe %>% drop_na(Right_to_Vote)
ggplot(europe, aes(x = Right_to_Vote, y = GDP)) + geom_boxplot(fill = '#C64955') + theme_minimal() + xlab("women's right to vote") + scale_x_discrete(labels = c("no","yes")) + ggtitle('Europe & Central Asia') + theme(plot.title = element_text(hjust = 0.5))

# North America, right to vote (Figure 9)

america = data %>% filter(Region == 'North America')
america$Right_to_Vote = as.factor(america$Right_to_Vote)
america = america %>% drop_na(Right_to_Vote)
ggplot(america, aes(x = Right_to_Vote, y = GDP)) + geom_boxplot(fill = '#533549', col = 'black') + theme_minimal() + xlab("women's right to vote") + scale_x_discrete(labels = c("no","yes")) + ggtitle('North America') + theme(plot.title = element_text(hjust = 0.5))

# Middle East & North Africa, right to vote (Figure 10)

east = data %>% filter(Region == 'Middle East & North Africa')
east$Right_to_Vote = as.factor(east$Right_to_Vote)
east = east %>% drop_na(Right_to_Vote)
ggplot(east, aes(x = Right_to_Vote, y = GDP)) + geom_boxplot(fill = '#9ab19a') + theme_minimal() + xlab("women's right to vote") + scale_x_discrete(labels = c("no","yes")) + ggtitle('Middle East & North Africa') + theme(plot.title = element_text(hjust = 0.5))
```

# Tests for correlations between variables

```{r}
#GDP_growth and regressors
cor.test(data$GDP_growth, data$Right_to_Vote) #-0.0405, significant
cor.test(data$GDP_growth, data$Share_in_Parliament) #-0.0040, significant

#Educ_Gender_Equality
cor.test(data$Educ_Gender_Equality, data$GDP_growth) #0.0914, significant
cor.test(data$Educ_Gender_Equality, data$Right_to_Vote) #0.0308, non-significant
cor.test(data$Educ_Gender_Equality, data$Share_in_Parliament) #0.2863, significant

#Female_Years_School
cor.test(data$Female_Years_School, data$GDP_growth) #0.1041, significant
cor.test(data$Female_Years_School, data$Right_to_Vote) #0.1305, significant
cor.test(data$Female_Years_School, data$Share_in_Parliament) #0.4439, significant

#Investments_In_Educatoin
cor.test(data$Investments, data$GDP_growth) #0.0357, non-significant
cor.test(data$Investments, data$Right_to_Vote) #0.02236, non-significant
cor.test(data$Investments, data$Share_in_Parliament) #0.2805, significant

#Democracy_Estimate
cor.test(data$Democracy_Estimate, data$GDP_growth) #0.0584, significant
cor.test(data$Democracy_Estimate, data$Right_to_Vote) #0.3213, significant
cor.test(data$Democracy_Estimate, data$Share_in_Parliament) #0.1861, significant

#Total_Fertility_Rate
cor.test(data$Total_Fertility_Rate, data$GDP_growth) #-0.0866, significant
cor.test(data$Total_Fertility_Rate, data$Right_to_Vote) #-0.2219, significant
cor.test(data$Total_Fertility_Rate, data$Share_in_Parliament) #-0.4001, significant
```

# Regression analysis

General models

```{r}
data$Right_to_Vote = as.factor(data$Right_to_Vote)
pdf_long = pdata.frame(data, index = c('Country', 'Year'))

# OSL models

ols_suffrage = plm(GDP_growth ~ Right_to_Vote + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, model = 'pooling')
ols_share = plm(GDP_growth ~ Share_in_Parliament + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + + Investments + Total_Fertility_Rate, data = pdf_long, model = 'pooling')

# TSLS models

tsls_suffrage = plm(GDP_growth ~ Right_to_Vote + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate | Anti_Patriarchy_Score + Democracy_Estimate  + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, effect = 'twoways', model = 'within')
tsls_share = plm(GDP_growth ~ Share_in_Parliament + Democracy_Estimate  + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate | Anti_Patriarchy_Score + Democracy_Estimate + Female_Years_School + Educ_Gender_Equality + Investments + Total_Fertility_Rate, data = pdf_long, effect = 'twoways', model = 'within')

stargazer(ols_suffrage, ols_share, tsls_suffrage, tsls_share, keep = c('Right_to_Vote', 'Share_in_Parliament', 'Democracy_Estimate', 'Educ_Gender_Equality', 'Female_Years_School', 'Investments', 'Total_Fertility_Rate', 'GDP_initial', 'Anti_Patriarchy_Score'), type = 'text', omit.stat = 'f', column.labels = c("OLS", "OLS", "TSLS", "TSLS"))

coeftest(ols_suffrage, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(ols_share, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(tsls_suffrage, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(tsls_share, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
```

Without obvious outliers

```{r}
no_outliers = data %>% filter(Country != 'Gabon') %>% filter(Country != 'Zimbabwe' | Year != 1994) %>% filter(Country != 'Zimbabwe' | Year != 1992)

no_outliers$Right_to_Vote = as.factor(no_outliers$Right_to_Vote)
pdf_long = pdata.frame(no_outliers, index = c('Country', 'Year'))

# OSL models

ols_suffrage = plm(GDP_growth ~ Right_to_Vote + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, model = 'pooling')
ols_share = plm(GDP_growth ~ Share_in_Parliament + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, model = 'pooling')

# TSLS models

tsls_suffrage = plm(GDP_growth ~ Right_to_Vote + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate | Anti_Patriarchy_Score + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, effect = 'individual', model = 'within')
tsls_share = plm(GDP_growth ~ Share_in_Parliament + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate | Anti_Patriarchy_Score + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, effect = 'twoways', model = 'within')

stargazer(ols_suffrage, ols_share, tsls_suffrage, tsls_share, keep = c('Right_to_Vote', 'Share_in_Parliament', 'Democracy_Estimate', 'Educ_Gender_Equality', 'Female_Years_School', 'Investments', 'Total_Fertility_Rate', 'GDP_initial', 'Anti_Patriarchy_Score'), type = 'text', omit.stat = 'f', column.labels = c("OLS", "OLS", "TSLS", "TSLS"))

coeftest(ols_suffrage, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(ols_share, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(tsls_suffrage, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(tsls_share, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
```

Without 0.995 and 0.005 quantiles in the variables of interest

```{r}
absolutely_no_outliers = data
absolutely_no_outliers$Right_to_Vote = as.numeric(absolutely_no_outliers$Right_to_Vote)

for(i in c(3,4,12)){
  
  Q1 = quantile(absolutely_no_outliers[,i], .005, na.rm = TRUE)
  Q3 = quantile(absolutely_no_outliers[,i], .995, na.rm = TRUE)
  
  absolutely_no_outliers = subset(absolutely_no_outliers, absolutely_no_outliers[,i] > Q1 | absolutely_no_outliers[,i] < Q3)
}


absolutely_no_outliers$Right_to_Vote = as.factor(absolutely_no_outliers$Right_to_Vote)
pdf_long = pdata.frame(absolutely_no_outliers, index = c('Country', 'Year'))

# OSL models

ols_suffrage = plm(GDP_growth ~ Right_to_Vote + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, model = 'pooling')
ols_share = plm(GDP_growth ~ Share_in_Parliament + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, model = 'pooling')

# TSLS models

tsls_suffrage = plm(GDP_growth ~ Right_to_Vote + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate | Anti_Patriarchy_Score + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, effect = 'twoways', model = 'within')
tsls_share = plm(GDP_growth ~ Share_in_Parliament + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate | Anti_Patriarchy_Score + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, effect = 'twoways', model = 'within')

stargazer(ols_suffrage, ols_share, tsls_suffrage, tsls_share, keep = c('Right_to_Vote', 'Share_in_Parliament', 'Democracy_Estimate', 'Educ_Gender_Equality', 'Female_Years_School', 'Investments', 'Total_Fertility_Rate', 'GDP_initial', 'Anti_Patriarchy_Score'), type = 'text', omit.stat = 'f', column.labels = c("OLS", "OLS", "TSLS", "TSLS"))


coeftest(ols_suffrage, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(ols_share, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(tsls_suffrage, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(tsls_share, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
```

Without Middle East & North Africa

```{r}
selected = data %>% filter(Region != 'Without Middle East & North Africa')

selected$Right_to_Vote = as.factor(selected$Right_to_Vote)
pdf_long = pdata.frame(selected, index = c('Country', 'Year'))

# OSL models

ols_suffrage = plm(GDP_growth ~ Right_to_Vote + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, model = 'pooling')
ols_share = plm(GDP_growth ~ Share_in_Parliament + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, model = 'pooling')

# TSLS models

tsls_suffrage = plm(GDP_growth ~ Right_to_Vote + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate | Anti_Patriarchy_Score + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, effect = 'individual', model = 'within')
tsls_share = plm(GDP_growth ~ Share_in_Parliament + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate | Anti_Patriarchy_Score + Democracy_Estimate + Educ_Gender_Equality + Female_Years_School + Investments + Total_Fertility_Rate, data = pdf_long, effect = 'twoways', model = 'within')

stargazer(ols_suffrage, ols_share, tsls_suffrage, tsls_share, keep = c('Right_to_Vote', 'Share_in_Parliament', 'Democracy_Estimate', 'Educ_Gender_Equality', 'Female_Years_School', 'Investments', 'Total_Fertility_Rate', 'GDP_initial', 'Anti_Patriarchy_Score'), type = 'text', omit.stat = 'f', column.labels = c("OLS", "OLS", "TSLS", "TSLS"))

coeftest(ols_suffrage, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(ols_share, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(tsls_suffrage, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
coeftest(tsls_share, df = Inf, vcov = vcovHC, type = 'HC0', method = "arellano", cluster = "group")
```
